Persistence in functional topology: q-tameness, local-connectivity and a correction to a theorem of Morse

### 2021-04-03 introduction.tex
Add section: historical remarks

\begin{itemize}
	\item Persistence homology, 
	\item applications in applied topology, application in theoretical topology, 
	\item barcode, stability, existence of barcode motivates q-tameness. 
	\item When can we ensure this condition? 
	\item Morse as a precursor. Geodesic study, natural next context: minimal surfaces, \item proposed condition for q-tameness, it's wrong, 
	\item a similar condition, assumed equivalent, we prove this implies it.
	\item We expect this paper to help the ongoing development of persistence methods in geometry.
\end{itemize}

@@

To achieve semi-continuity, Morse assumes all spaces in his filtrations to be compact and uses Vietoris homology, which agrees with \v{C}ech homology by Dowker's Theorem \cite{Dowker.1952}. This implies semi-continuity by a general result stating that, for field coefficients, \v{C}ech homology commutes with inverse limits for compact Hausdorff spaces \cite[Theorem VIII.3.6 and Theorem X.3.1]{MR0050886}. To achieve q-tameness, Morse wants to use the following condition, given e.g.\@ in \cite{Morse.1940}.

\begin{defi}
	Let $M$ be a metric space and $f \colon M \to \R$ a function. The space $M$ is called \emph{locally-$f$-connected} if for each $p\in M$, $e>0$ and $d\in\mathbb{N}_0$ there exists $\delta\in(0,e)$ such that the map 
	\[
	\tilde{H}_d(f_{\leq f(p)+\delta}\cap B_{\delta}(p)\hookrightarrow f_{\leq f(p)+e}\cap B_e(p);\mathbb{Z})
	\]
	is 0.
\end{defi}

Here, $\tilde{H}_{d}$ denotes reduced $d$-dimensional singular homology, $f_{\leq t}$ denotes the sublevel set of $f$ at $t\in\R$, and $B_{\delta}(p)$ denotes the open ball with radius $\delta$ around the point $p$. Morse then states the claim below as \cite[Theorem 6.3]{Morse.1940}.

\begin{claim}
	Let $M$ be a metric space and $f\colon M\to\R$ a function such that $M$ is locally-$f$-connected and all sublevel sets are compact. Then $\CH_d(f_{\leq\bullet};\mathbb{F})$ is q-tame for any $d$ and any field $\mathbb{F}$.
\end{claim}

Here, $\CH_{d}$ denotes $d$-dimensional Vietoris or \v{C}ech homology and $f_{\leq\bullet}$ denotes the sublevel set filtration of $f$. He does not prove the statement in the above reference, but cites \cite[Theorem 6.1]{Morse.1938}. There, Morse presents a supposed proof of the claim by constructing finite bases for the sets $\im \CH_d(f_{\leq s}\hookrightarrow f_{\leq t};\mathbb{F})$. As it turns out, the proof is not correct: We use the Hawaiian earring to construct a counterexample to this claim in Section \ref{sec:counterexample}.

In \cite{Morse.1937}, Morse gives a slightly stronger definition of local-$f$-connectedness, which we will define under a different name to distinguish it from the weaker notion.

\begin{defi} 
	Let $M$ be a metric space and $f\colon M\to\R$ a function. The space $M$ is called \emph{locally-$f$-connected} if for each $p \in M$, $e > 0$ and $d \in \mathbb{N}_0$ there exists $\delta \in (0,e)$ such that for all $c \geq f(p)$ the map 
	\[
	\tilde{H}_d(f_{\leq c+\delta} \cap B_{\delta}(p) \hookrightarrow f_{\leq c+e} \cap B_e(p); \mathbb{Z})
	\]
	is 0.
\end{defi}

Clearly, strong local-$f$-connectedness implies local-$f$-connectedness.
Morse also presents the following claim as \cite[Theorem 9.2]{Morse.1937}, but states immediately thereafter that the proof "while not difficult involves too great detail to be presented here".

\begin{claim}
	Let $M$ be a metric space and $f \colon M \to \R$ a function such that $M$ is strongly locally-$f$-connected and all sublevel sets are compact.
	Then $\CH_d(f_{\leq\bullet};\mathbb{F})$ is q-tame for any $d$ and any field $\mathbb{F}$.
\end{claim}

We devote Section \ref{sec:theorem} to its proof.

>>>>

\begin{itemize}
	\item Persistence homology, 
	\item applications in applied topology, application in theoretical topology, 
	\item barcode, stability, existence of barcode motivates q-tameness. 
	\item When can we ensure this condition? 
	\item Morse as a precursor. Geodesic study, natural next context: minimal surfaces, \item proposed condition for q-tameness, it's wrong, 
	\item a similar condition, assumed equivalent, we prove this implies it.
	\item We expect this paper to help the ongoing development of persistence methods in geometry.
	\item Floer theory and simplectic geometry.
\end{itemize}

@@

### 2021-04-03 theorem.tex
Add section: historical remarks

\begin{defi}
	A sublevel filtration $(X,f)$ is said to be \textit{strongly} $\HLC$ if for any $\epsilon > 0$, $x \in X$, and $V$ an open neighborhood of $x$, there is a $\delta \in (0, \epsilon)$ and a neighborhood $U$ of $x$ such that $U \subseteq V$ and
	\begin{equation*}
	X_{\leq f(x) + \delta + c} \cap U \to X_{f(x) + \epsilon + c} \cap V
	\end{equation*}
	is $\HT$ for every $c \geq 0$.
\end{defi}

>>>>

\begin{defi} \label{defi:strong_local_connectedness}
	A sublevel filtration $(X,f)$ is said to be \textit{strongly} $\HLC$ if for any $\epsilon > 0$, $x \in X$, and $V$ an open neighborhood of $x$, there is a $\delta \in (0, \epsilon)$ and a neighborhood $U$ of $x$ such that $U \subseteq V$ and
	\begin{equation*}
	X_{\leq f(x) + \delta + c} \cap U \to X_{f(x) + \epsilon + c} \cap V
	\end{equation*}
	is $\HT$ for every $c \geq 0$.
\end{defi}
